Word Learning as Network Growth: a Cross-linguistic Analysis

### 2018-01-25 cogsci/cogsci.Rmd
modif in Cogci

A crucial question that naturally emerges from these analyses is how does network connectivity influence word learning?  @steyvers2005 suggested that the effect of connectivity is the consequence of how the lexical network gets constructed in the child's mind. According to this explanation, known as Preferential Attachment (PAT), highly connected words in the child's lexicon tend to "attract" more words over time, in a rich-get-richer scenario [@barabasi99]. In other words, what predicts word learning is *internal* connectivity in the child's early lexicon. In contrast, @hills2009 found that what biases the learning is not the connectivity in the child's internal lexicon but, rather, *external* connectivity in the learning environment. They called this alternative explanation Preferential Acquisition (PAC).  Figure \ref{fig:growth} shows an illustration of both growth scenarios with the same simplified network.

@@

Studies that investigated lexical network growth focused on semantic networks using English data only [@steyvers2005; @hills2009; @hills2010].  The novelty of the current study is that it investigates whether phonological networks, like semantic networks, grow by PAC, or if they rather grow by PAT. It also provides a systematic comparison of network growth scenarios in  the phonological and the semantic domains and assesses their relative contribution to the learning process. Moreover, it tests the generality of the findings across eight languages.

@@

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.width=7, fig.height=3, fig.align = "center", strip.white=TRUE, fig.cap = "\\label{fig:data_all}Age of acquisition in the end-state network as predicted by the degree in this network. Results are shown in each language for the phonological network (top) and the semantic network (bottom). Each point is a word, with lines indicating linear model fits."}
plot_data_all
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

plot_dist <- ggplot(data = degreeDist,  aes(x=x, y=y, col=dimension))+
  facet_grid(. ~ language)+
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme_bw()+
  theme(aspect.ratio = 1)+
  xlab("degree") +ylab("probability")

@@

We also analysed the global network degree distribution. PAT is known to generate networks with a power-law degree distribution [@barabasi99]. Thus, if the networks display this property, it means that PAT cannot be ruled out *a priori* as a potential growth mechanism. In contrast, if there is strong evidence against a power law distribution, then PAT (but not PAC) can be ruled out.  The distributions are shown in Figure \ref{fig:degree_distribution}. Overall, the results do not provide strong evidence against a power-law distribution. Moreover, the degree to which the distribution approximates a power law varies across dimension and languages. For a more conclusive test, we fit explicit growth models to the data (next section).
<!--
```{r }

@@

To test the network growth scenarios, we fit different growth models to the data. We proceeded as follows. We calculated the probability that a word $w_i$, with a growth value $d_i$ would enter the lexicon at a given month, using a softmax function: 
\begin{equation}
 p(w_i)= \frac{e^{\beta d_i}}{\sum_j e^{\beta d_j} }
\end{equation}
Where $\beta$ is the fitting parameter. A positive value of $\beta$ means that words with higher growth values $d_i$ are acquired first, and a negative value means that words with lower growth values are acquired first. As explained in Figure \ref{fig:growth}, each candidate word $w_i$ is characterized with a growth values $d_i$ that depends on the growth scenario tested.  In PAT, the growth value of a given word is equal to the average degree of the existing nodes that it would attach to. In PAC, the growth value of a word is equal to its degree in the entire network.  The normalization includes all words that could be learned at that month.

@@

Above we evaluated the network growth scenarios individually. As a next step, we analysed their relative contribution to the learning process. This was done through adding more fitting parameters to the model, that is, by substituting $\beta d_i$ in formula (1) with: $$\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}$$ where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and PhonoPAC. Using the same fitting technique, we obtained the values shown in Figure \ref{fig:pred_ind}. In term of growth scenarios, we found that PAC dominates the learning. In particular, though we found previously that phonological PAT predicted part of the learning in some languages when tested individually, here PAT appears to lose its predictive power when pitted against phonological PAC. In terms of domain, both phonological and semantic networks appear to contribute to learning, although the phonological network appears to be stronger and more reliable across languages. In summary, the findings show that both semantic and phonological networks grow primarily by PAC, and that, generally speaking, semantic and phonological networks both contribute to the learning process.

@@

Since PAT was uninformative, we dropped it from this analysis. Thus, we no longer needed to fit the growth model month-by-month as in the previous section. In fact, word utilities in the case of PAC are fixed, they do not depend on previously learned words. A more direct way to assess and compare the contribution of PAC in relation to other factors is through conducting linear regressions, where connectivity in the learning environment (at both the phonological and semantic level), frequency and length predict the age of acquisition.

@@

We conducted two analyses. We fit a linear regression for each language, and we fit a linear mixed-effect to all the data pooled across languages, with language as a random effect. Figure \ref{fig:regressions_img} shows the coefficient estimate for each predictor in each language, and figure \ref{fig:regressions_all_img} shows the coefficient estimates for all languages combined (all predictors were centered and scaled). The findings were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition.  Word length predicts learning in some languages such as Croatian and Norwegian, but not in others (including English). It remains, however, a significant predictor in the global model. As for the factors of interest, i.e., semantic and phonological connectivity, we also found cross-linguistic differences. The phonological connectivity contributes to learning in languages such as Croatian, English and Russian, whereas semantic connectivity contributes to learning in Turkish, Spanish and to some extent in Croatian, but interestingly not in English\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included both frequency and length as covariates. Another important difference is the number of words tested. Our study uses a larger set of nouns.}. Despite this cross-linguistic variation, both phonological and semantic connectivity remain significant predictors in the global model.

@@

Finally, this study shares a number of limitations with previous studies using similar datasets. <!--For instance, we used normative age of acquisition. However, individual children may follow different learning trajectories. The use of longitudinal data would allow us to assess how the aggregate behavior relates to these individual trajectories [e.g., typical vs. late talkers, @beckage2011]. Moreover,--> In particular, the results provide correlational but not causal evidence. Thus, the conclusions of this study  require parallel evidence, especially from controlled behavioral experiments.

>>>>

While most studies focued on the static properties of the lexical network, a few have proposed to account for the underlying developmental process. @steyvers2005 suggested that the effect of connectivity is the consequence of how the lexical network gets constructed in the child's mind. According to this explanation, known as Preferential Attachment (PAT), highly connected words in the child's lexicon tend to "attract" more words over time, in a rich-get-richer scenario [@barabasi99]. In other words, what predicts word learning is *internal* connectivity in the child's early lexicon. In contrast, @hills2009 suggested that what biases the learning is not the connectivity in the child's internal lexicon but, rather, *external* connectivity in the learning environment. They called this alternative explanation Preferential Acquisition (PAC).  Figure \ref{fig:growth} shows an illustration of both growth scenarios with the same simplified network.

@@

Studies that investigated lexical network growth focused on semantic networks using English data only [@steyvers2005; @hills2009; @hills2010].  The novelty of the current study is that it investigates whether phonological networks, like semantic networks, grow by PAC, or if they rather grow by PAT. It also provides a systematic comparison of both network growth scenarios in the phonological and the semantic domains and assesses their relative contribution to the learning process. Moreover, it tests the generality of the findings across eight languages.

@@

```{r all_data, fig.env = "figure*", fig.pos = "h", fig.width=7, fig.height=3, fig.align = "center", strip.white=TRUE, fig.cap = "\\label{fig:data_all}Age of acquisition in the global network as predicted by the degree in this network. Results are shown in each language for the phonological network (top) and the semantic network (bottom). Each point is a word, with lines indicating linear model fits."}
plot_data_all
#img <- png::readPNG("figs/data_all.png")
#grid::grid.raster(img)

@@

plot_dist <- ggplot(data = degreeDist,  aes(x=x, y=y, col=dimension))+
  facet_grid(. ~ language)+
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme_bw()+
  theme(aspect.ratio = 0.7)+
  xlab("degree") +ylab("probability")

@@

We also analysed the global network degree distribution. PAT is known to generate networks with a power-law degree distribution [@barabasi99]. If the networks display this property, this fact would strongly suggests a PAT-like generative process. <!--cannot be ruled out *a priori* as a potential growth mechanism. In contrast, if there is strong evidence against a power law distribution, then PAT (but not PAC) can be ruled out--> The results are shown in Figure \ref{fig:degree_distribution}. Overall, the extent to which the degree distribution approximates a power law varies across dimensions and languages. For a more direct test of the developmental process, we fit explicit growth models to the data (see next section).
<!--
```{r }

@@

To test the network growth scenarios, we fit different growth models to the data. We proceeded as follows. We calculated the probability that a word $w_i$, with a growth value $d_i$ would enter the lexicon at a given month, using a softmax function: 
\begin{equation}
 p(w_i)= \frac{e^{\beta d_i}}{\sum_j e^{\beta d_j} }
\end{equation}
Where $\beta$ is the fitting parameter. A positive value of $\beta$ means that words with higher growth values $d_i$ are acquired first, and a negative value means that words with lower growth values are acquired first. As explained in Figure \ref{fig:growth}, each candidate word $w_i$ is characterized with a growth value $d_i$ that depends on the growth scenario tested.  In PAT, the growth value of a given word is equal to the average degree of the existing nodes that it would attach to. In PAC, the growth value of a word is equal to its degree in the entire network.  The normalization includes all words that could be learned at that month.

@@

Above we evaluated the network growth scenarios individually. As a next step, we analysed their relative contribution to the learning process. This was done through adding more fitting parameters to the model, that is, by substituting $\beta d_i$ in formula (1) with: $$\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}$$ where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and PhonoPAC. Using the same fitting technique, we obtained the values shown in Figure \ref{fig:pred_ind}. In term of growth scenarios, we found that PAC dominates the learning. <!--In particular, though we found previously that phonological PAT predicted part of the learning in some languages when tested individually, here PAT appears to lose its predictive power when pitted against phonological PAC--> In terms of domain, both phonological and semantic networks appear to contribute to the lexical growth, but the phonological network appears to be stronger and more consistent across languages. In summary, the findings show that both semantic and phonological networks grow primarily by PAC, and that, generally speaking, semantic and phonological networks both contribute to the learning process.

@@

Since PAT was uninformative, we dropped it from this analysis. Thus, we no longer needed to fit the growth model month-by-month as in the previous section. In fact, word utilities (i.e., growth values) in the case of PAC are fixed, they do not depend on previously learned words. A more direct way to assess and compare the contribution of PAC in relation to other factors is through conducting linear regressions, where connectivity in the learning environment (at both the phonological and semantic level), frequency and length predict the age of acquisition.

@@

We conducted two analyses. We fit a linear regression for each language, and we fit a linear mixed-effect to all the data pooled across languages, with language as a random effect. Figure \ref{fig:regressions_img} shows the coefficient estimate for each predictor in each language, and figure \ref{fig:regressions_all_img} shows the coefficient estimates for all languages combined (all predictors were centered and scaled). The findings were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition.  Word length predicts learning in some languages such as Croatian and Norwegian, but not in others (including English). It remains, however, a significant predictor in the global model. As for the factors of interest, i.e., semantic and phonological connectivity, we also found cross-linguistic differences. The phonological connectivity contributes to learning in languages such as Croatian, English and Russian, whereas semantic connectivity contributes to learning in Turkish, Spanish and to some extent in Croatian, but interestingly not in English\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included both frequency and length as covariates. Another important difference is the number of words tested: our study uses a larger set of nouns.}. Despite this cross-linguistic variation, both phonological and semantic connectivity remain significant predictors in the global model.

@@

Finally, this study shares a number of limitations with previous studies using similar research strategy and dataset. <!--For instance, we used normative age of acquisition. However, individual children may follow different learning trajectories. The use of longitudinal data would allow us to assess how the aggregate behavior relates to these individual trajectories [e.g., typical vs. late talkers, @beckage2011]. Moreover,--> In particular, the results provide correlational but not causal evidence. Thus, the conclusions of this study  require parallel evidence, especially from controlled behavioral experiments.

### 2018-01-25 cogsci/cogsci.tex
modif in Cogci

A crucial question that naturally emerges from these analyses is how
does network connectivity influence word learning? Steyvers \& Tenenbaum
(2005) suggested that the effect of connectivity is the consequence of
how the lexical network gets constructed in the child's mind. According
to this explanation, known as Preferential Attachment (PAT), highly
connected words in the child's lexicon tend to ``attract'' more words
over time, in a rich-get-richer scenario (Barabasi \& Albert, 1999). In
other words, what predicts word learning is \emph{internal} connectivity
in the child's early lexicon. In contrast, Hills et al. (2009) found
that what biases the learning is not the connectivity in the child's
internal lexicon but, rather, \emph{external} connectivity in the
learning environment. They called this alternative explanation
Preferential Acquisition (PAC). Figure \ref{fig:growth} shows an
illustration of both growth scenarios with the same simplified network.

@@

Studies that investigated lexical network growth focused on semantic
networks using English data only (Hills et al., 2010, 2009; Steyvers \&
Tenenbaum, 2005). The novelty of the current study is that it
investigates whether phonological networks, like semantic networks, grow
by PAC, or if they rather grow by PAT. It also provides a systematic
comparison of network growth scenarios in the phonological and the
semantic domains and assesses their relative contribution to the
learning process. Moreover, it tests the generality of the findings
across eight languages.

@@

\caption{\label{fig:data_all}Age of acquisition in the end-state network as predicted by the degree in this network. Results are shown in each language for the phonological network (top) and the semantic network (bottom). Each point is a word, with lines indicating linear model fits.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

@@

We also analysed the global network degree distribution. PAT is known to
generate networks with a power-law degree distribution (Barabasi \&
Albert, 1999). Thus, if the networks display this property, it means
that PAT cannot be ruled out \emph{a priori} as a potential growth
mechanism. In contrast, if there is strong evidence against a power law
distribution, then PAT (but not PAC) can be ruled out. The distributions
are shown in Figure \ref{fig:degree_distribution}. Overall, the results
do not provide strong evidence against a power-law distribution.
Moreover, the degree to which the distribution approximates a power law
varies across dimension and languages. For a more conclusive test, we
fit explicit growth models to the data (next section).

@@

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher growth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. As explained in Figure \ref{fig:growth}, each candidate
word \(w_i\) is characterized with a growth values \(d_i\) that depends
on the growth scenario tested. In PAT, the growth value of a given word
is equal to the average degree of the existing nodes that it would
attach to. In PAC, the growth value of a word is equal to its degree in
the entire network. The normalization includes all words that could be
learned at that month.

@@

Above we evaluated the network growth scenarios individually. As a next
step, we analysed their relative contribution to the learning process.
This was done through adding more fitting parameters to the model, that
is, by substituting \(\beta d_i\) in formula (1) with:
\[\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}\]
where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and
PhonoPAC. Using the same fitting technique, we obtained the values shown
in Figure \ref{fig:pred_ind}. In term of growth scenarios, we found that
PAC dominates the learning. In particular, though we found previously
that phonological PAT predicted part of the learning in some languages
when tested individually, here PAT appears to lose its predictive power
when pitted against phonological PAC. In terms of domain, both
phonological and semantic networks appear to contribute to learning,
although the phonological network appears to be stronger and more
reliable across languages. In summary, the findings show that both
semantic and phonological networks grow primarily by PAC, and that,
generally speaking, semantic and phonological networks both contribute
to the learning process.

@@

Finally, this study shares a number of limitations with previous studies
using similar datasets. In particular, the results provide correlational
but not causal evidence. Thus, the conclusions of this study require
parallel evidence, especially from controlled behavioral experiments.

>>>>

While most studies focued on the static properties of the lexical
network, a few have proposed to account for the underlying developmental
process. Steyvers \& Tenenbaum (2005) suggested that the effect of
connectivity is the consequence of how the lexical network gets
constructed in the child's mind. According to this explanation, known as
Preferential Attachment (PAT), highly connected words in the child's
lexicon tend to ``attract'' more words over time, in a rich-get-richer
scenario (Barabasi \& Albert, 1999). In other words, what predicts word
learning is \emph{internal} connectivity in the child's early lexicon.
In contrast, Hills et al. (2009) suggested that what biases the learning
is not the connectivity in the child's internal lexicon but, rather,
\emph{external} connectivity in the learning environment. They called
this alternative explanation Preferential Acquisition (PAC). Figure
\ref{fig:growth} shows an illustration of both growth scenarios with the
same simplified network.

@@

Studies that investigated lexical network growth focused on semantic
networks using English data only (Hills et al., 2010, 2009; Steyvers \&
Tenenbaum, 2005). The novelty of the current study is that it
investigates whether phonological networks, like semantic networks, grow
by PAC, or if they rather grow by PAT. It also provides a systematic
comparison of both network growth scenarios in the phonological and the
semantic domains and assesses their relative contribution to the
learning process. Moreover, it tests the generality of the findings
across eight languages.

@@

\caption{\label{fig:data_all}Age of acquisition in the global network as predicted by the degree in this network. Results are shown in each language for the phonological network (top) and the semantic network (bottom). Each point is a word, with lines indicating linear model fits.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

@@

We also analysed the global network degree distribution. PAT is known to
generate networks with a power-law degree distribution (Barabasi \&
Albert, 1999). If the networks display this property, this fact would
strongly suggests a PAT-like generative process. The results are shown
in Figure \ref{fig:degree_distribution}. Overall, the extent to which
the degree distribution approximates a power law varies across
dimensions and languages. For a more direct test of the developmental
process, we fit explicit growth models to the data (see next section).

@@

Where \(\beta\) is the fitting parameter. A positive value of \(\beta\)
means that words with higher growth values \(d_i\) are acquired first,
and a negative value means that words with lower growth values are
acquired first. As explained in Figure \ref{fig:growth}, each candidate
word \(w_i\) is characterized with a growth value \(d_i\) that depends
on the growth scenario tested. In PAT, the growth value of a given word
is equal to the average degree of the existing nodes that it would
attach to. In PAC, the growth value of a word is equal to its degree in
the entire network. The normalization includes all words that could be
learned at that month.

@@

Above we evaluated the network growth scenarios individually. As a next
step, we analysed their relative contribution to the learning process.
This was done through adding more fitting parameters to the model, that
is, by substituting \(\beta d_i\) in formula (1) with:
\[\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}\]
where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and
PhonoPAC. Using the same fitting technique, we obtained the values shown
in Figure \ref{fig:pred_ind}. In term of growth scenarios, we found that
PAC dominates the learning. In terms of domain, both phonological and
semantic networks appear to contribute to the lexical growth, but the
phonological network appears to be stronger and more consistent across
languages. In summary, the findings show that both semantic and
phonological networks grow primarily by PAC, and that, generally
speaking, semantic and phonological networks both contribute to the
learning process.

@@

Finally, this study shares a number of limitations with previous studies
using similar research strategy and dataset. In particular, the results
provide correlational but not causal evidence. Thus, the conclusions of
this study require parallel evidence, especially from controlled
behavioral experiments.
